---
title: "cRazy-coders-module7"
author: "cRazy-coders"
date: "2022-11-14"
output: html_document
bibliography: BIOL3140.bib
---

```{r, include=FALSE}
library(tidyverse)
library(MuMIn)
library(knitr)
library(dplyr)
library(ggplot2)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
The primary goal of this project is to explore the force-length relationship in muscles and specifically how this relationship shifts when muscles are fatigued. The mechanics of the force-length relationship in muscles is well studied, as described by @huxley2004fifty, it is the number of cross-bridges formed on a fiber level that mediate the force by a muscle along with the sliding filaments that explain the force-length relationship of sarcomeres. A maximally activated single fiber is experiencing maximum filament overlap to do so. The sliding filament theory is based on the idea that length changes in sarcomeres, fibers, and muscles happen by relative sliding of the stiff myofilaments, actin, and myosin, within a sarcomere.
 
In order to study such a relationship, we will be conducting controlled and fatigued experiments using the human upper limb muscles by lifting a heavy book bag. This will cause the muscles to experience eccentric fatigue, in which the muscles are doing negative work. By undertaking maximum voluntary contractions (MVCs), our group will produce isometric force-angle curves for human forearm flexors. We will then use this data to compare at what angle maximum isometric force for fatigued and non-fatigued forearm flexors. 

# Methods
Before capturing MVC data, a ganiometer was created using a circular piece of cardboard. The angle marks were added to the circle to complete the ganiometer. Then the ganiometer would be adhered to the elbow of each team member during data collection at approximately the lateral epicondyle of the humerus. This would allow us to measure the angle formed at the elbow joint between the forearm and upper arm during our experiment. The MVC data will be collected through a portable Arduino-based force-data acquisition system that we designed and constructed. Isometric force MVCs will be collected over 12 arm angles (45-157.5 degrees at 11). With the elbow rested on a sturdy surface, control data will be collected for each angle. At each angle, the person will flex their elbow as intensely as possible to produce the MVC. Once control data has been collected, one will need to eccentrically fatigue their forearm flexors through isokinetic controlled drops of a heavy book bag weighing ⅓ of their maximum isometric force obtained during the control experiments. After three minutes of the isokinetic controlled drops, the forearm flexors should be sufficiently fatigued. With fatigued forearm flexors, repeat the steps taken to obtain the control data. 

Following data collection, the force-length relationship will be analyzed in two ways. First through the isometric force-angle curves, and second through comparing the angles at which the maximum isometric force occurred between the control and experimental data. The maximum force for each recording was determined and normalized against the control and experimental data. The normalized values then were analyzed to determine if any shifts occurred in maximum force in relation to the angle. Next, the differences in maximum force between the two data sets were found and the mean shifts were calculated. 


# Results
```{r, Read files and making tibble}
f <- list.files(path="./Project 8 data", full.names = T)
dat.1 <- list()

for(i in f){
  metadata <- unlist(strsplit(i,"_"))
  sub <- metadata[2]
  ang <- as.numeric(metadata[3])
  exp <- gsub(".csv", " ", metadata[4])
  dat.1[[i]] <- read_delim(i,delim = " ", col_names = c("Reading", "Force", "Unit"))%>%
    mutate(
      subject = sub,
      angle = ang,
      experiment = exp
)
}

dat <- do.call(rbind, dat.1)
dat.maxrecording <- dat%>%
  group_by(subject,angle,experiment)%>%
  summarize(max_force_recording= max(abs(Force)))
  print(dat.maxrecording)

dat.maxexp <- dat.maxrecording%>%
  group_by(subject,experiment)%>%
  summarize(max_force_exp=max(max_force_recording))
  print(dat.maxexp)  

dat.norm <- dat.maxrecording%>%
  left_join(dat.maxexp)%>%
  print(dat.norm)

dat.norm2 <- dat.norm%>%
  group_by(angle, subject, experiment)%>%
  summarize(max_force_normalized=max_force_recording/max_force_exp)
  print(dat.norm2)
  
```

The normalized data of the force length relationship is shown below. It is evident that the class data does not follow the normal expected force length relationship.

```{r, Cleaning up of tables and plotting}
dat.2 <- dat.norm2%>%
  na.omit()

dat.2%>%
  ggplot(aes(angle,max_force_normalized,col=experiment))+geom_point()
```
***Figure 1. Normalized data of force-length relationship in control and fatigued muscles ***

Below, we start looking for the model that best fits the data by doing an AIC analysis.
```{r, AICs}
AICs <- dat.norm2%>%
  group_by(subject,experiment)%>%
  summarize(
    m2=AICc(lm(max_force_normalized~poly(angle,2))), #second order
    m3=AICc(lm(max_force_normalized~poly(angle,3))), #third order
    m4=AICc(lm(max_force_normalized~poly(angle,4))) #fourth order
  )%>%
  pivot_longer(m2:m4,names_to="model",values_to="AICc")%>%
  print()

```

```{r, Table of fits of different models, message = FALSE}
x.pred <- seq(45,157.5,length.out = 1000)

fits <- dat.norm2%>%
  group_by(subject,experiment)%>%
  summarize(
    m2=predict(lm(max_force_normalized~poly(angle,2)),newdata=data.frame(angle=x.pred)), #second order
    m3=predict(lm(max_force_normalized~poly(angle,3)),newdata=data.frame(angle=x.pred)), #third order
    m4=predict(lm(max_force_normalized~poly(angle,4)),newdata=data.frame(angle=x.pred)) #fourth order
  )%>%
   pivot_longer(m2:m4,names_to="model")%>%
   group_by(subject,experiment,model)%>%
   summarize(theta_max=x.pred[which.max(value)])%>%
   print()

```

```{r, Finding best model, message = FALSE}
best.models <- fits%>%
  left_join(AICs)%>%
  group_by(subject,experiment)%>%
  mutate(best=AICc==min(AICc))%>%
  filter(best==TRUE)%>%
  dplyr::select(-best)%>%
  print()

```

```{r, Anova}
anova(lm(theta_max~experiment,best.models))
```

```{r, Finding theta shift}
dat.3 <- best.models%>%
  pivot_wider(id_cols=subject,names_from = experiment, values_from=theta_max)%>%
  print()

dat.3 <- dat.3 %>% mutate(shift= `fatigue `-`control `)%>% #Why do the variable names change like this, it took forever to figure out
  ungroup() %>%
  na.omit()%>% 
  summarise(mean.shift=mean(shift),se.shift=sd(shift)/sqrt(length(shift))) %>% 
  print()


```

# Discussion
A graph of the class wide data for the force angle relationship of the human forearm flexors deviates from a normal force-length relationship as there is an increase in maximum force, whereas a bell shaped distribution would normally be seen. As found in @macnaughton2006reports, there is a rightward shift in the force length relationship after sufficient fatigue. As the muscle was subjected to fatigue, the active force used at a longer muscle length increased from the previous trial that was done prior to the fatigue protocol. 

A difference between the θmax value is shown between the control muscle and the fatigued muscle, as the values from the control were around 140 degrees, and the fatigued were around 157 degrees. An ANOVA analysis on this data reported a p value of 0.04718, which is less that 0.05 therefore signifying a significant relationship between the post-MVC fatigued muscle and control muscle force-angle relationship.

# Author Contributions
Eugene - Results
Katherine - Introduction
Phoebe - Methods
Sabrina - Discussion

