---
title: "cRazy-coders-module5"
author: "cRazy-coders"
date: "2022-10-05"
output: html_document
bibliography: BIOL2140.bib
---

```{r, include=FALSE}
library(tidyverse)
library(knitr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction

In this project we will be exploring whether body temperature vary in human subjects according to activity level. As suggested by a study by @refinetti1999relationship , in many observed mammals an increase in metabolic heat production derived from the increase in activity may be responsible for the body temperature rhythm. Data was collected regarding the locomotive activity and body temperature of 8 mammalian species, four of them nocturnal (flying squirrel, golden hamster, fat-tailed gerbil, and Mongolian gerbil) and four of them diurnal (Chilean degu, thirteen-lined ground squirrel, Richardson ground squirrel, and tree shrew), in order to study the relationship between these two rhythms. The plotted results for the flying squirrel showed daily rhythmicity in both temperature and activity, where increases in activity were often accompanied by corresponding increases in body temperature. In all four mammalian species studied, the results similarly showed the daily rhythms of body temperature and locomotive activity being very closely synchronized. Therefore, a increase in activity has been shown to lead to an increase in body temperature and vice versa. We will be investigating this principle in our study.


# Methods

As mentioned in the introduction, the relationship between body temperature in humans with activity level is being explored in this study. In order to determine the relationship between the two, each human subject had their body temperature measured from their armpit while calmly sitting for 30s. Then, the subjects were instructed to perform jump and jacks for 1 minute. After exerting themselves for 1 minute, the body temperatures of the subjects were once again recorded from their armpits. Each subject had their data stored in individual files which were then combined and analyzed in R Markdown. 

# Results
```{r, include=FALSE}
library(tidyverse)
dat.f <- list.files("Data", full.names=TRUE,pattern = "Rest|Active.csv")
dat.l <- list()

for(i in dat.f){
  met.dat<- unlist(strsplit(i,"_"))
  who <- basename(met.dat[1])
  activity <- gsub(".csv","",met.dat[2])
  dat.l[[i]]<- read_csv(i)%>%
    mutate(who=who,activity=activity)
}
dat <- do.call(rbind,dat.l)
```
```{r, echo=FALSE}
dat%>%
  ggplot() + geom_boxplot(aes(x=activity,y=degC)) + facet_grid(.~who) + labs(caption="Box-Plot Showing Activity Type for Each Group Member (x-axis) vs Body Temperature in Celcius (y-axis)") + theme(plot.caption=element_text(hjust=0.5))
```

# Discussion

From our data we observed that our body temperatures were lower after a minute of exercising than when resting. Our findings contradict congruent findings from @refinetti1999relationship, however, this may be due to the length of exercise and differences between species. The subjects exercised for the short 1 minute period after rest, which cannot be compared to a days worth of locomotive activity. Humans also sweat to balance the heat produced during exercise, shown in a study by @kenney2017restoration, and thus the human body temperature will drop due to loss of heat from sweat. This could explain the results we observed from our experiment, as our body temperatures dropped after exercise. 

# Author Contributions

Katherine - Introduction, References, Results Revisions  
Eugene - Methods, Device Assembly, Data Collection  
Phoebe - Data + Results, References, Revision + Edits  
Sabrina - Discussion, Device Assembly, Code Revisions

<!-- for comments -->
<style>
div.comment pre { background-color:#ffbbbb; }
div.comment pre.r { background-color:#ffbbbb; }
div.comment {color: red;}
</style>

<div class="comment">
Top notch stuff. Looks great. But, you don't need to annotate your references VVV

15/15

</div>
# References

@refinetti1999relationship studied locomotive activity vs temperature in four mammals. 

@kenney2017restoration studied sweat's role in balancing temperature during exercise.
